---
title: MPhil Thesis Data
author: Devjyoti Chakrabarti 
---

```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(ggplot2)
library(scales)
library(readr)
library(dplyr)
library(tidyr)
```

## India Data

### Consumer Price Index

Consumer Price Index, General Index, All India. Unclear if it is seasonally adjusted.\
**Source:** [MOSPI](https://cpi.mospi.gov.in/Default1.aspx)

```{r, message=FALSE, echo=FALSE, warning=FALSE}
# Read the data
data <- read_csv("Data/India Data/CPI_Index.csv")

# Convert observation_date to Date format (dd/mm/yyyy)
data$DATE <- as.Date(data$DATE, format="%d/%m/%Y")

# Create the plot with aesthetic improvements
ggplot(data, aes(x = DATE, y = Combined)) +
  geom_line(color = "#1f77b4", size = 1.2) +  # Change to a more appealing blue line
  labs(
    title = "CPI Index",
    x = "Date",  # Label for the x-axis
    y = "CPI"  # Label for the y-axis
  ) +
  scale_x_date(
    breaks = seq(min(data$DATE), max(data$DATE), by = "2 year"),  # Show labels every 6 months
    labels = scales::date_format("%b %Y")  # Format labels as Month Year (e.g., Nov 2024)
  ) +
  theme_minimal() +  # Minimal theme for a clean look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"),  # Rotate and style x-axis labels
    axis.text.y = element_text(size = 12, color = "black"),  # Style y-axis labels
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),  # Bold x-axis title
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),  # Bold y-axis title
    plot.title = element_text(size = 16, color = "black", face = "bold", hjust = 0.5),  # Bold and centered title
    legend.position = "none",  # No legend as it's a single variable
    axis.line = element_line(color = "black", size = 0.5),  # Black axis lines
    panel.grid.major = element_line(color = "grey90", size = 0.5),  # Light grey grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    panel.background = element_rect(fill = "white", color = "white")  # White background
  )
```

### Real Gross Domestic Product

Millions of Domestic Currency, Not Seasonally Adjusted, Quarterly <br> **Source:** [FRED](https://fred.stlouisfed.org/series/NGDPRNSAXDCINQ)

```{r, message=FALSE, echo=FALSE}
data <- read_csv("Data/India Data/NGDPRNSAXDCINQ.csv")

# Convert observation_date to Date format (dd/mm/yyyy)
data$DATE <- as.Date(data$DATE, format="%d/%m/%Y")

# Create the plot with aesthetic improvements
ggplot(data, aes(x = DATE, y = NGDPRNSAXDCINQ)) +
  geom_line(color = "#1f77b4", size = 1.2) +  # Change to a more appealing blue line
  labs(
    title = "Real Gross Domestic Product (Millions Rupee)",
    x = "Date",  # Label for the x-axis
    y = "Real GDP"  # Label for the y-axis
  ) +
  scale_x_date(
    breaks = seq(min(data$DATE), max(data$DATE), by = "2 year"),  # Show labels every 6 months
    labels = scales::date_format("%b %Y")  # Format labels as Month Year (e.g., Nov 2024)
  ) +
  theme_minimal() +  # Minimal theme for a clean look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"),  # Rotate and style x-axis labels
    axis.text.y = element_text(size = 12, color = "black"),  # Style y-axis labels
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),  # Bold x-axis title
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),  # Bold y-axis title
    plot.title = element_text(size = 16, color = "black", face = "bold", hjust = 0.5),  # Bold and centered title
    legend.position = "none",  # No legend as it's a single variable
    axis.line = element_line(color = "black", size = 0.5),  # Black axis lines
    panel.grid.major = element_line(color = "grey90", size = 0.5),  # Light grey grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    panel.background = element_rect(fill = "white", color = "white")  # White background
  )
```

### Repo Rate and Reverse Repo Rate

Percent <br> **Source:** [RBI](https://www.rbi.org.in/scripts/PublicationsView.aspx?id=22517)

```{r, message=FALSE, echo=FALSE}
# Load the data
data <- read_csv("Data/India Data/Money_Rates.csv")

# Ensure the DATE column is properly formatted
data$DATE <- as.Date(data$DATE, format="%d/%m/%Y")

# Create the plot
ggplot(data) +
  geom_line(aes(x = DATE, y = Repo_Rate, color = "Repo Rate"), size = 1.2) +
  geom_line(aes(x = DATE, y = Reverse_Repo_Rate, color = "Reverse Repo Rate"), size = 1.2) +
  labs(
    title = "Repo and Reverse Repo Rates Over Time",
    x = "Date",
    y = "Rate (%)",
    color = "Legend"
  ) +
  scale_x_date(
    breaks = seq(min(data$DATE), max(data$DATE), by = "1 year"),
    labels = date_format("%b %Y")
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    plot.title = element_text(size = 16, color = "black", face = "bold", hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    axis.line = element_line(color = "black", size = 0.5),
    panel.grid.major = element_line(color = "grey90", size = 0.5),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = "white")
  )
```

### Industrial Production Index

Monthly indices of industrial production as per use-based classification <br> **Source:** [MOSPI](https://mospi.gov.in/iip)

```{r, message=FALSE, echo=FALSE}
# Load the data
data <- read_csv("Data/India Data/IIP.csv")

# Convert 'DATE' column to Date format (if not already in the correct format)
data$DATE <- as.Date(data$DATE, format = "%d/%m/%Y")

# Reshape the data from wide to long format for ggplot
data_long <- data %>%
  pivot_longer(
    cols = c(Primary_goods, Capital_goods, Intermediate_goods, 
             Infrastructure_or_construction_goods, Consumer_durables, Consumer_non_durables),
    names_to = "Category",
    values_to = "Value"
  )

# Create the plot
ggplot(data_long, aes(x = DATE, y = Value, color = Category)) +
  geom_line(size = 1.2) +  # Plot all lines with different colors
  labs(
    title = "Industrial Production Index by Category",
    x = "Date",  # Label for the x-axis
    y = "Industrial Production Index",  # Label for the y-axis
    color = "Category"  # Legend title
  ) +
  scale_x_date(
    breaks = seq(min(data$DATE), max(data$DATE), by = "1 year"),  # Show labels every year
    labels = scales::date_format("%b %Y")  # Format labels as Month Year (e.g., Nov 2024)
  ) +
  theme_minimal() +  # Minimal theme for a clean look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"),  # Rotate and style x-axis labels
    axis.text.y = element_text(size = 12, color = "black"),  # Style y-axis labels
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),  # Bold x-axis title
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),  # Bold y-axis title
    plot.title = element_text(size = 16, color = "black", face = "bold", hjust = 0.5),  # Bold and centered title
    legend.position = "bottom",  # Place the legend at the bottom
    legend.title = element_text(size = 12, face = "bold"),  # Style legend title
    legend.text = element_text(size = 10),  # Style legend text
    axis.line = element_line(color = "black", size = 0.5),  # Black axis lines
    panel.grid.major = element_line(color = "grey90", size = 0.5),  # Light grey grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    panel.background = element_rect(fill = "white", color = "white")  # White background
  )
```
